### 2024-09-28 article.qmd
Remove duplicate coordinates in wildfire dataset

This paper aims to delineate the wildland-urban interface in the sparsely populated, limited resourced Atlantic-Boreal Europe. Due to its demographic and biophysical differences to southern Europe, as well as to other fire-prone regions, current WUI definitions to suit the needs of southern countries are not viable. Therefore, common WUI mapping assumptions, and how well the resulting WUI maps (thresholds) capture the reality of wildfire in northern Europe, are scoped.

@@

This work has been financed by XXX

>>>>

This paper aims to delineate the wildland-urban interface in the sparsely populated, limited-resourced Scandinavian peninsula (excl. Finland). Common WUI mapping assumptions and how different thresholds capture the reality of building ignition from wildfire in this region are evaluated. We show that dedicated fuel maps capture areas at risk slightly better than vegetation maps, although the choice of map per se is less important than the selected limit for possible ember transport. The commonly used 6.17 buildings/km^2^ threshold for building density fails to capture most wildfire incidents that have led to building ignition.

@@

N. Fernandez-Anez acknowledges The Research Council of Norway, Project 301569: Building Design for At-risk Groups (BUILDER) for supporting her contribution. F. Vermina Plathner was supported by FORMAS \[grant number 2021-02396\]. E. Pastor acknowledges the Spanish Ministry of Science, Innovation and Universities, Project PID2023-150607OB-I00.

### 2024-09-28 article_codes.qmd
Remove duplicate coordinates in wildfire dataset

Area of wild vegetation by county (@tbl-summary-table-areas): we calculate the number vegetation cells (i.e., grid centroid is inside the polygon), and calculate the area based on the raster resolution.

@@

Number of buildings in WUI

```{r}
#| label: tbl-buildings-wui
#| tbl-cap: Number of buildings in each WUI classification

c(build_counts, wui_r) |> 
  as_tibble() |>   
  mutate(sum = ifelse(is.na(sum), 0, sum)) |> 
  drop_na() |> 
  pivot_longer(col = -sum, names_to = "name") |> 
  group_by(name, value) |>
  summarise(buildings = sum(sum)) |> 
  ungroup() |> 
  pivot_wider(names_from = name, 
              values_from = buildings) |> 
  gt(rowname_col =  "value") |> 
  grand_summary_rows(
    fns = list(Total = ~sum(.))
  ) |> 
   cols_label(
    corine = md("**corine**"),
    effis = md("**effis**"),
    firEUrisk ~ md("**firEUrisk**"),
    national ~ md("**national**")
  )

@@

wf_se <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
  janitor::clean_names() |> 
  filter(datetime >= 2016) |> 
  st_as_sf(coords = c("ecoord", "nccord"),
           crs =  3006) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "SE")

wf <- rbind(wf_no, wf_se)

ggplot() +
  geom_sf(data = region_sf, fill = "lightgrey") +
  geom_sf(data = country_sf, fill = NA) +
  geom_sf(data = wf, col = "red", size = 0.4) +
  theme_void()

@@

We have analysed all wildfires that have affected buildings from 2016 to 2023 (i.e., `r dim(wf)[1]`; `r dim(filter(wf, country == "NO"))[1]` in Norway and \``r dim(filter(wf, country == "SE"))[1]` in Sweden). The can be obtained from XX (Norway) and XXX (Sweden).

@@

brks <-  c(0,  1, 3, 6, 10, 20, 50, 100, 500, 1700)
wf_build_density |> 
  mutate(bd_brks = cut(bd, 
                       brks,
                       include.lowest = TRUE,
                       right = FALSE,
                       dig.lab = 4)) |> 
  group_by(bd_brks) |> 
  summarise(n = n(), 
            perc = n() / nrow(wf_build_density) * 100) |> 
  mutate(cs = cumsum(perc)) |> 
  gt() |> 
  fmt_number(columns = c("perc", "cs"),
             decimals = 1) |> 
  tab_spanner(label = md("**Wildfires**"),
              columns = c(n, perc, cs)) |> 
  cols_label(
    bd_brks = md(html("**Building**<br>**density**")),
    n  = md("**Number**"),
    perc = md("**Percentage [%]**"),
    cs = md("**Cumulative [%]**")
  )

>>>>

Area of wild vegetation by county (@tbl-summary-table-areas): we calculate the number of vegetation cells (i.e., grid centroid is inside the polygon), and calculate the area based on the raster resolution.

@@

Total area of WUI (by type of vegetation):

```{r}
#| label: area-wui-type

wui_r |>  
  as_tibble() |> 
  drop_na() |> 
  pivot_longer(col = everything(),
               names_to = "name") |> 
  group_by(name, value) |>
  summarise(n = n()) |> 
  ungroup() |> 
  mutate(area = n * (0.25 * 0.25)) |> #km2
  pivot_wider(names_from = name, 
              values_from = c(n, area)) |> 
  select(value, starts_with("area")) |> 
  gt() |>
  fmt_number(decimals = 0) |> 
  cols_label(
    value = "",
    area_corine = md("**corine**"),
    area_effis = md("**effis**"),
    area_firEUrisk ~ md("**firEUrisk**"),
    area_national ~ md("**national**")
  ) |> 
  tab_options(
    table.font.size = 9,
    row_group.background.color = "lightgrey",
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

Number of buildings in WUI
```{r}
#| label: tbl-buildings-wui
#| tbl-cap: Number of buildings in each WUI classification

c(build_counts, wui_r) |> 
  as_tibble() |>   
  mutate(sum = ifelse(is.na(sum), 0, sum)) |> 
  drop_na() |> 
  pivot_longer(col = -sum, names_to = "name") |> 
  group_by(name, value) |>
  summarise(buildings = sum(sum)) |> 
  ungroup() |> 
  pivot_wider(names_from = name, 
              values_from = buildings) |> 
  gt(rowname_col =  "value") |> 
  grand_summary_rows(
    fns = list(Total = ~sum(.))
  ) |> 
   cols_label(
    corine = md("**corine**"),
    effis = md("**effis**"),
    firEUrisk ~ md("**firEUrisk**"),
    national ~ md("**national**")
  ) |>
  tab_options(
    table.font.size = 9,
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )

@@

# Infrastructures affected by wildfires in SE
wf_se_inf <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
  janitor::clean_names() |> 
  filter(datetime >= 2016) |> 
  st_as_sf(coords = c("ecoord", "nccord"),
           crs =  3006) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "SE")

# Wildfires in SE (unique coordinates - one wildfire can affect several buildings)
wf_se <- distinct(wf_se_inf)

# Final dataset (NO + SE)
wf <- rbind(wf_no, wf_se)

# Map
ggplot() +
  geom_sf(data = region_sf, fill = "#FAFAFA") +
  geom_sf(data = country_sf, fill = NA) +
  geom_sf(data = wf, col = "red", size = 0.4) +
  theme_void()

@@

We have analysed all wildfires that have affected buildings from 2016 to 2023 (i.e., `r dim(wf)[1]`; `r dim(filter(wf, country == "NO"))[1]` in Norway and `r dim(filter(wf, country == "SE"))[1]` in Sweden). The can be obtained from the *Norwegian Directorate for Civil Protection (DSB)*  (Norway) and the *Swedish Civil Contingencies Agency (MSB)* (Sweden). [Note: They were not uploaded in our GitHub repository]

@@

brks <-  c(0,  1, 3, 6, 10, 20, 50, 100, 500, 1700)
wf_build_density |> 
  mutate(bd_brks = cut(bd, 
                       brks,
                       include.lowest = TRUE,
                       right = FALSE,
                       dig.lab = 4)) |> 
  group_by(bd_brks) |> 
  summarise(n = n(), 
            perc = n() / nrow(wf_build_density) * 100) |> 
  mutate(cs = cumsum(perc)) |> 
  gt() |> 
  fmt_number(columns = c("perc", "cs"),
             decimals = 1) |> 
  tab_spanner(label = md("**Wildfires**"),
              columns = c(n, perc, cs)) |> 
  cols_label(
    bd_brks = md(html("**Building**<br>**density**")),
    n  = md("**Number**"),
    perc = md("**Percentage [%]**"),
    cs = md("**Cumulative [%]**")
  ) |> 
  tab_options(
    table.font.size = 9,
    row_group.padding = px(1),
    row_group.font.weight = "bold",
    data_row.padding = px(1)
  )
